MSDS_6372_Project2

### 2022-07-27 MSDS_6372_Project2.Rmd
interim update

#### Loading the necessary R libraries for the analysis
```{r message = FALSE}
# Load the necessary libraries
library(knitr)
library(rmarkdown)
# From class
library(GGally)
library(epitools)
library(MASS)
library(tidyverse)
library(car)
library(caret)
library(glmnet)
library(ROCR)
#Dasta set
library(aplore3)

@@

#library(ggpubr)
#library(tidyr)
#library(plyr)
#library(ggthemes)
#library(e1071)
#library(class)
#library(stringr)
#library(sjPlot)
#library(data.table)
#library(reshape2)
#library(corrplot)
#library(naivebayes)
#library(egg)
#library(rworldmap)
#library(Hmisc)
#library(DataExplorer)
#library(selectiveInference)
#library(dlookr)
```
```{r message = FALSE}
# Turn off scientific notation
options(scipen = 100, digits = 4)
```

@@

### Check VIF
```{r, vif orig}
vif(multi_var.log)
```

### Remove the multicollinearity
```{r, all variables manual wo multicollinearity}
multi_var.log.vif<-glm(fracture~priorfrac+age+weight+height+premeno+momfrac+armassist+smoke+raterisk+fracscore+bonemed+
                     bonemed_fu+bonetreat,family=binomial(link='logit'),data=train)
multi_var.log.vif.sum = summary(multi_var.log.vif)
multi_var.log.vif.sum$coefficients
```

### re-Check VIF
```{r, vif}
vif(multi_var.log.vif)
```

### Observations:
* Fitting all variables to the logistic regression model, it shows that only 'bonmed', 'bonemed_fu' and 'bonetreat' are statistically significant. 
* We observed multicollinearity between 'weight' and 'bmi'. We removed the 'bmi' variable and re-run the logistic regression model.
* As a result the following variables are statistically significant:
1. height
2. bonemed
3. bonemed_fu
4. bonetreat

@@

```{r, lasso feature selection}
grid=10^seq(10,-2, length =100)
lasso.mod = glmnet(x, y, family = binomial(link='logit'), alpha=1, lambda=grid)
cv.out=cv.glmnet(x,y,family = "binomial", alpha=1)
plot(cv.out)

bestlambda = cv.out$lambda.min
coef(lasso.mod,s=bestlambda)

# Final LASSO model
lasso.mod.final=glmnet(x,y,family = binomial(link='logit'), alpha=1, type.measure = "class", lambda=bestlambda)
```

@@

```{r, y.test}
y.test = model.matrix(fracture~., test)[,-c(1,17)]
```
#####################################################################################
#                                       LASSO                                       #
#####################################################################################
```{r, LASSO Predict}
cutoff.lasso<-0.43
# Predicting on the final LASSO model TRAINing data
lasso.mod.final.pred.train <- predict(lasso.mod.final, newx = x, type = "response")

# Confusion matrix
class.lasso.final.train<-factor(ifelse(lasso.mod.final.pred.train>cutoff.lasso,"Yes","No"),levels=c("No","Yes"))

#Confusion Matrix for LASSO
print("Confusion matrix for LASSO TRAINING with 0.5 cutoff")
confusionMatrix(table(class.lasso.final.train,train$fracture))

################ Misclassification rate train ######################################
#cross.table.lasso.train = table(table(class.lasso.final.train,train$fracture))
#MCR_lasso.train = (cross.table.lasso.train[2]+cross.table.lasso.train[3])/dim(train)[1]
misClasificError.lasso.train = mean(class.lasso.final.train != train$fracture)
print(paste('Misclassification Rate for lasso on training set: ', misClasificError.lasso.train))
#####################################################################################################
# Predicting on the final LASSO model TEST data
lasso.mod.final.pred.test <- predict(lasso.mod.final, newx = y.test, type = "response")

# Confusion matrix
class.lasso.final.test<-factor(ifelse(lasso.mod.final.pred.test>cutoff.lasso,"Yes","No"),levels=c("No","Yes"))

#Confusion Matrix for LASSO
print("Confusion matrix for LASSO TEST with 0.5 cutoff")
confusionMatrix(table(class.lasso.final.test,test$fracture))

################ Misclassification rate test ######################################
#cross.table.lasso.test = table(class.lasso.final.test,test$fracture)
#MCR_lasso.test = (cross.table.lasso.test[2]+cross.table.lasso.test[3])/dim(test)[1]
#print(paste('Misclassification Rate for LASSO selection on test set: ', MCR_lasso.test))
misClasificError.lasso.test = mean(class.lasso.final.test != test$fracture)
print(paste('Misclassification Rate for lasso on test set: ', misClasificError.lasso.test))

################ LASSO ROC Curve ######################################
library(ROCR)
lasso.roc.test = prediction(lasso.mod.final.pred.test, test$fracture,label.ordering=c("No","Yes"))
roc.lasso.test = performance(lasso.roc.test, measure = "tpr", x.measure = "fpr")
plot(roc.lasso.test,colorize = TRUE)
abline(a=0, b= 1)
```

@@

@@

#####################################################################################
#                                       Stepwise                                    #
#####################################################################################
```{r, Stepwise Predict}
cutoff.step = 0.42
# Predicting on the final Stepwise model TRAINing data
stepwise.mod.final.pred.train <- predict(step.log, newdata = train, type = "response")

# Confusion matrix
class.stepwise.final.train<-factor(ifelse(stepwise.mod.final.pred.train>cutoff.step,"Yes","No"),levels=c("No","Yes"))

#Confusion Matrix for Stepwise
print("Confusion matrix for Stepwise TRAINING with 0.5 cutoff")
confusionMatrix(table(class.stepwise.final.train,train$fracture))

################ Misclassification rate train ######################################
#cross.table.stepwise.train = table(class.stepwise.final.train,train$fracture)
#MCR_stepwise.train = (cross.table.stepwise.train[2]+cross.table.stepwise.train[3])/dim(train)[1]
#print(paste('Misclassification Rate for stepwise selection on training set: ', MCR_stepwise.train))

misClasificError.stepwise.train = mean(class.stepwise.final.train != train$fracture)
print(paste('Misclassification Rate for stepwise on train set: ', misClasificError.stepwise.train))

#####################################################################################################
# Predicting on the final Stepwise model TEST data
stepwise.mod.final.pred.test <- predict(step.log, newdata = test, type = "response")

@@

#Confusion Matrix for Stepwise
print("Confusion matrix for Stepwise TEST with 0.5 cutoff")
confusionMatrix(table(class.stepwise.final.test,test$fracture))

################ Misclassification rate test ######################################
#cross.table.stepwise.test = table(class.stepwise.final.test,test$fracture)
#MCR_stepwise.test = (cross.table.stepwise.test[2]+cross.table.stepwise.test[3])/dim(test)[1]
#print(paste('Misclassification Rate for stepwise selection on test set: ', MCR_stepwise.test))

misClasificError.stepwise.test = mean(class.stepwise.final.test != test$fracture)
print(paste('Misclassification Rate for stepwise on test set: ', misClasificError.stepwise.test))

@@

################ Stepwise ROC Curve ######################################
library(ROCR)
stepwise.roc.test = prediction(stepwise.mod.final.pred.test, test$fracture,label.ordering=c("No","Yes"))
roc.stepwise.test = performance(stepwise.roc.test, measure = "tpr", x.measure = "fpr")
plot(roc.stepwise.test,colorize = TRUE)
abline(a=0, b= 1)
```

@@

MLogReg = train(
  fracture ~ age + height + bmi + priorfrac,
  data = train,
  method = "glm",
  family = "binomial",
  trControl = cv,
  metric = "logLoss")

@@

# Predicting
MLogReg.pred.train = predict(MLogReg, train)
MLogReg.pred.test = predict(MLogReg, test)

#Confusion Matrix for manual
confusionMatrix(table(MLogReg.pred.train,train$fracture))
confusionMatrix(table(MLogReg.pred.test,test$fracture))

################ Misclassification rate train ######################################
cross.table.manual.train = table(MLogReg.pred.train,train$fracture)
MCR_manual.train = (cross.table.manual.train[2]+cross.table.manual.train[3])/dim(train)[1]
print(paste('Misclassification Rate for manual selection on training set: ', MCR_manual.train))

################ Misclassification rate test ######################################
cross.table.manual.test = table(MLogReg.pred.test,test$fracture)
MCR_manual.test = (cross.table.manual.test[2]+cross.table.manual.test[3])/dim(test)[1]
print(paste('Misclassification Rate for manual selection on test set: ', MCR_manual.test))

################ Manual ROC Curve ######################################
#library(ROCR)
#manual.roc.test = prediction(as.numeric(MLogReg.pred.test), as.numeric(test$fracture), label.ordering=c("No","Yes"))
#roc.manual.test = performance(manual.roc.test, measure = "tpr", x.measure = "fpr")
#plot(roc.manual.test,colorize = TRUE)
#abline(a=0, b= 1)
```

@@

```{r, interaction}
cutoff.manual = 0.5
LogReg.Interaction<-glm(fracture~age + height + bmi + priorfrac, family=binomial(link='logit'))
LogReg.Interaction.pred.train <- predict(LogReg.Interaction, newdata = train, type = "response")

Anova(LogReg.Interaction,type=3)

# Confusion matrix
class.LogReg.Interaction.train<-factor(ifelse(LogReg.Interaction.pred.train>cutoff.manual,"Yes","No"),levels=c("No","Yes"))

#Confusion Matrix for interaction
print("Confusion matrix for Interaction TRAINING with 0.5 cutoff")
confusionMatrix(table(class.LogReg.Interaction.train,train$fracture))

################ Misclassification rate train ######################################
cross.table.interaction.train = table(class.LogReg.Interaction.train,train$fracture)
MCR_interaction.train = (cross.table.interaction.train[2]+cross.table.interaction.train[3])/dim(train)[1]
print(paste('Misclassification Rate for stepwise selection on training set: ', MCR_interaction.train))
```

>>>>

#### Loading the necessary R libraries for the analysis
```{r message = FALSE}
# Load the necessary libraries
library(knitr)
library(rmarkdown)
# From class
library(GGally)
library(epitools)
library(MASS)
library(tidyverse)
library(car)
library(caret)
library(glmnet)
library(ROCR)
library(pROC)
#Dasta set
library(aplore3)
library(ResourceSelection)

@@

#library(ggpubr)
#library(tidyr)
#library(plyr)
#library(ggthemes)
library(e1071)
#library(class)
#library(stringr)
#library(sjPlot)
#library(data.table)
#library(reshape2)
#library(corrplot)
#library(naivebayes)
#library(egg)
#library(rworldmap)
#library(Hmisc)
#library(DataExplorer)
#library(selectiveInference)
#library(dlookr)
```
```{r message = FALSE}
# Turn off scientific notation
options(scipen = 100, digits = 4)
```

@@

### Observations:
* Fitting all variables to the logistic regression model, it shows that only 'rateriskGreater', 'bonmed', 'bonemed_fu' and 'bonetreat' are statistically significant.

@@

```{r, lasso feature selection}
# 5-fold cross validation
cv.lasso <- trainControl(
  method = "repeatedcv", 
  number = 5,
  repeats = 10,
  savePredictions = TRUE,
  summaryFunction=mnLogLoss,
  classProbs = TRUE
)

lasso.logreg.mod = train(
  fracture ~  priorfrac+age+weight+height+bmi+premeno+momfrac+armassist+smoke+raterisk+fracscore+bonemed+bonemed_fu+bonetreat,
  data = train,
  method = "glmnet",
  family = "binomial",
  trControl = cv.lasso,
  metric = "logLoss")

plot(lasso.logreg.mod)
coef(lasso.logreg.mod$finalModel,lasso.logreg.mod$finalModel$lambdaOpt)

# Final LASSO model refit with glm
lasso.mod.final=glm(fracture ~  priorfrac+age+height+bmi+premeno+momfrac+armassist+smoke+raterisk+fracscore+bonemed+bonemed_fu+bonetreat, 
                    family = binomial(link='logit'), data = train)

# ODD ratios for interpretation
exp(cbind("Odds ratio" = coef(lasso.mod.final), confint.default(lasso.mod.final, level = 0.95)))
```

## Observations and interpretations of the LASSO coeffficients:

@@

```{r, y.test}
y.test = model.matrix(fracture~., test)[,-c(1,17)]
```
#####################################################################################
#                                       LASSO                                       #
#####################################################################################
```{r, LASSO Predict}

# Predicting on the final LASSO model TRAINing data
lasso.mod.final.pred.train <- predict(lasso.mod.final, newdata = train, type = "response")

# Predicting on the final LASSO model TEST data
lasso.mod.final.pred.test <- predict(lasso.mod.final, newdata = test, type = "response")

################ LASSO ROC Curve ######################################

lasso.roc.test = prediction(lasso.mod.final.pred.test, test$fracture,label.ordering=c("No","Yes"))
roc.lasso.test = performance(lasso.roc.test, measure = "tpr", x.measure = "fpr")
plot(roc.lasso.test,colorize = TRUE)
abline(a=0, b= 1)

AUC = auc(test$fracture, lasso.mod.final.pred.test)
print(paste("Area Under the Cuve: ", AUC))
print(paste("LASSO AIC Score: ", AIC(lasso.mod.final)))

#alternative approach
lasso.roc<-roc(response=test$fracture,predictor=lasso.mod.final.pred.test,levels=c("No","Yes"))
plot(lasso.roc,print.thres="best") #This graph is nice because the x axis is plotted in terms of specificity rather than FPR
auc(lasso.roc)

cutoff.lasso = 0.212

# Confusion matrix train
class.lasso.final.train = factor(ifelse(lasso.mod.final.pred.train>cutoff.lasso,"Yes","No"),levels=c("No","Yes"))

#Confusion Matrix for LASSO
print("Confusion matrix for LASSO TRAINING")
confusionMatrix(table(class.lasso.final.train,train$fracture), positive = "Yes")

################ Misclassification rate train ######################################
#cross.table.lasso.train = table(table(class.lasso.final.train,train$fracture))
#MCR_lasso.train = (cross.table.lasso.train[2]+cross.table.lasso.train[3])/dim(train)[1]
misClasificError.lasso.train = mean(class.lasso.final.train != train$fracture)
print(paste('Misclassification Rate for lasso on training set: ', misClasificError.lasso.train))
#####################################################################################################

# Confusion matrix test
class.lasso.final.test<-factor(ifelse(lasso.mod.final.pred.test>cutoff.lasso,"Yes","No"),levels=c("No","Yes"))

#Confusion Matrix for LASSO
print("Confusion matrix for LASSO TEST with 0.5 cutoff")
confusionMatrix(table(class.lasso.final.test,test$fracture), positive = "Yes")

################ Misclassification rate test ######################################
#cross.table.lasso.test = table(class.lasso.final.test,test$fracture)
#MCR_lasso.test = (cross.table.lasso.test[2]+cross.table.lasso.test[3])/dim(test)[1]
#print(paste('Misclassification Rate for LASSO selection on test set: ', MCR_lasso.test))
misClasificError.lasso.test = mean(class.lasso.final.test != test$fracture)
print(paste('Misclassification Rate for lasso on test set: ', misClasificError.lasso.test))
```

@@

* Sensitivity: The probability that the model predicts a positive outcome for an observation when indeed the outcome is positive. This is also called the “true positive rate.”
* Specificity: The probability that the model predicts a negative outcome for an observation when indeed the outcome is negative. This is also called the “true negative rate.”

@@

#####################################################################################
#                                       Stepwise                                    #
#####################################################################################
```{r, Stepwise Predict}

# Predicting on the final Stepwise model TRAINing data
stepwise.mod.final.pred.train <- predict(step.log, newdata = train, type = "response")

# Predicting on the final Stepwise model TEST data
stepwise.mod.final.pred.test <- predict(step.log, newdata = test, type = "response")

################ Stepwise ROC Curve ######################################
library(ROCR)
stepwise.roc.test = prediction(stepwise.mod.final.pred.test, test$fracture,label.ordering=c("No","Yes"))
roc.stepwise.test = performance(stepwise.roc.test, measure = "tpr", x.measure = "fpr")
plot(roc.stepwise.test,colorize = TRUE)
abline(a=0, b= 1)

AUC = auc(test$fracture, stepwise.mod.final.pred.test)
print(paste("Area Under the Cuve: ", AUC))
print(paste("Stepwise AIC Score: ", AIC(step.log)))

#alternative approach
step.roc<-roc(response=test$fracture,predictor=stepwise.mod.final.pred.test,levels=c("No","Yes"))
plot(step.roc,print.thres="best") #This graph is nice because the x axis is plotted in terms of specificity rather than FPR

cutoff.step = 0.159

# Confusion matrix stepwise train
class.stepwise.final.train<-factor(ifelse(stepwise.mod.final.pred.train>cutoff.step,"Yes","No"),levels=c("No","Yes"))

#Confusion Matrix for Stepwise train
print("Confusion matrix for Stepwise TRAINING with 0.5 cutoff")
confusionMatrix(table(class.stepwise.final.train,train$fracture), positive = "Yes")

#####################################################################################################

@@

#Confusion Matrix for Stepwise
print("Confusion matrix for Stepwise TEST with 0.5 cutoff")
confusionMatrix(table(class.stepwise.final.test,test$fracture), positive = "Yes")

#misClasificError.stepwise.test = mean(class.stepwise.final.test != test$fracture)
#print(paste('Misclassification Rate for stepwise on test set: ', misClasificError.stepwise.test))

@@

```

@@

MLogReg = train(
  fracture ~  age + bonemed + bonemed_fu + bonetreat,
  data = train,
  method = "glm",
  family = "binomial",
  trControl = cv,
  metric = "logLoss")

@@

# Predicting
MLogReg.pred.train = predict(MLogReg, train, type = 'prob')
MLogReg.pred.test = predict(MLogReg, test, type = 'prob')

AUC = auc(test$fracture, MLogReg.pred.test$Yes)
print(paste("Area Under the Cuve: ", AUC))
print(paste("Manual model AIC Score: ", MLogReg$finalModel$aic))

################ Manual ROC Curve ######################################
manual.roc<-roc(response=test$fracture,predictor=MLogReg.pred.test$Yes,levels=c("No","Yes"))
plot(manual.roc,print.thres="best")

cutoff.manual = 0.231

#Confusion Matrix for manual model
class.manual.train<-ifelse(MLogReg.pred.train$Yes > cutoff.manual,"Yes","No")
class.manual.train<-factor(class.manual.train)

class.manual.test<-ifelse(MLogReg.pred.test$Yes > cutoff.manual,"Yes","No")
class.manual.test<-factor(class.manual.test)

confusionMatrix(table(class.manual.train,train$fracture), positive = "Yes")
confusionMatrix(table(class.manual.test,test$fracture), positive = "Yes")

```

@@

```{r, interaction}
cv.ii <- trainControl(
  method = "repeatedcv", 
  number = 5,
  repeats = 10,
  savePredictions = TRUE,
  summaryFunction=mnLogLoss,
  classProbs = TRUE
)

MLogReg.ii = train(
  fracture ~  age + height + bonemed_fu + bonetreat + bmi*bonemed_fu + bmi*bonemed + I(age^2) + I(height^3),
  data = train,
  method = "glm",
  family = "binomial",
  trControl = cv,
  metric = "logLoss")

### Hypothesis testing
summary(MLogReg.ii$finalModel)

anova(MLogReg.ii$finalModel, test="Chisq")

hoslem.test(MLogReg.ii$finalModel$y,fitted(MLogReg.ii))

# Predicting
MLogReg.ii.pred.train = predict(MLogReg.ii, train, type = 'prob')
MLogReg.ii.pred.test = predict(MLogReg.ii, test, type = 'prob')

AUC = auc(test$fracture, MLogReg.ii.pred.test$Yes)
print(paste("Area Under the Cuve: ", AUC))
print(paste("Manual model with interactions and polynomial terms AIC Score: ", MLogReg.ii$finalModel$aic))

################ Manual ROC Curve ######################################
manual.roc<-roc(response=test$fracture,predictor=MLogReg.ii.pred.test$Yes,levels=c("No","Yes"))
plot(manual.roc,print.thres="best")

cutoff.manual = 0.23

#Confusion Matrix for manual model
class.manual.train.ii<-ifelse(MLogReg.ii.pred.train$Yes > cutoff.manual,"Yes","No")
class.manual.train.ii<-factor(class.manual.train)

class.manual.test.ii<-ifelse(MLogReg.ii.pred.test$Yes > cutoff.manual,"Yes","No")
class.manual.test.ii<-factor(class.manual.test)

confusionMatrix(table(class.manual.train.ii,train$fracture), positive = "Yes")
confusionMatrix(table(class.manual.test.ii,test$fracture), positive = "Yes")

### Visualize VIF
#MLogReg_ii_VIF = vif(MLogReg.ii$finalModel)
#barplot(MLogReg_ii_VIF, main = 'VIF Values (Custom Logistic Regression with interactions', horiz = TRUE, col="blue", xlim = c(0,12))
#abline(v=10, col="red")

```
